Hydrodynamics and rheology of weakly adhesive vesicle doublets

### 2019-01-30 vesicle_code/docs/adhesion/adhesion.tex
more chnages in intro by yny

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by excess area (reduced volume), viscosity contrast
between interior and exterior fluids, and shear rate of the imposed
far-field fluid flow. In addition a vesicle with a rigid particle inside
is also investigated as a biological mimic of a eukaryotic cell with a
nucleus that occupies nearly half of the intracellular
volume~\cite{Veerapaneni2011_PRL}. Small-deformation analysis shows that
a vesicle tank-treads in a planar shear flow for low viscosity contrast
and shear rate. At high viscosity contrast the tank-tread (TT) dynamics
transitions to tumbling (TB)~\cite{Misbah2006_PRL, Vlahovska2007_PRE},
and this leads to a transition in effective shear viscosity in the
vesicle suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also
validated by direct numerical
simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,
ZabuskySegreDeschamps2011_PoF}.  Between TT and TB vesicle
hydrodynamics, a breathing (tremble) is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume  and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
Collective hydrodynamics of vesicles is dictated by the individual
vesicle-vesicle interactions.  
In a quiescent flow,
vesicle-vesicle adhesion leads to formation of vesicle doublets \cite{Ziherl2007_PRL,ZiherlSvetina2007_PNAS} or clusters \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
As a model for red blood cell (RBC) aggregates, an idealized model for adhesive vesicle-vesicle interactions gives rise to 
vesicle configurations similar to those observed in experiments of fibrinogen-induced RBC aggregates \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}. 
Using the L-J potential for point-point interaction between two vesicle membranes without the detailed molecular details of
adhesive interactions between RBCs, Flormann {\it et al.} found the contact zone to buckle under strong adhesion \cite{FlormannAouane2017_SciReports}. 
A direct consequence of the buckling is the
sigmoidal shape of the adhesion region, which is also observed in RBC doublets and explained in a slightly different model \cite{ZiherlSvetina2007_PNAS}. 
It remains unknown how adhesive interactions between vesicles might lead to different vesicle hydrodynamics that results in
different rheological properties of a vesicle suspension. Such studies will provide useful insight to the rheology of blood flow and how to use nano solutes to control the hypertension
by adjusting the adhesive interactions between RBCs.
%

In numerical simulations of  a vesicle suspension, the adhesive interaction between two vesicles is often ignored \cite{Veerapaneni2009_JCP,RahimianVeerapaneniBiros2010_JCP} 
for numerical convenience
because a smaller time step is often needed to resolve the dynamics of the lubrication thin film between two membranes with adhesion than without.
However, adhesive interactions  between RBCs lead to different RBC clusters that are expected to alter the rheological properties of the RBC suspension \cite{NeuMeiselman2002_BJ,SvetinaZiherl2008_Bioelectrochemistry}.
Hydrodynamic simulations of adhesive vesicles in a quiescent flow have revealed physical insights to the equilibrium shapes of RBC aggregates in both experiments 
\cite{FlormannAouane2017_SciReports} and theory \cite{ZiherlSvetina2007_PNAS}. The main goal in this work is to investigate the hydrodynamics of adhesive vesicles in flowing
conditions common in microfluidic vesicle flow, such as a planar shear flow or an extensional flow.

%
The electrostatic nature of lipid molecules (a hydrophobic tail and a hydrophilic head
with an electric dipole) complicates the interactions between a lipid
bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces,PerutkovaFrank-Bertoncelij2013_CSB} 
or a solid (such as a glass substrate
or a nano particle).  In the absence of an external electric field and ions in the solvents, 
it is reported that 
%such complicated electrostatic component in 
the adhesive interactions between two membranes can be well approximated by the 
L-J-type potential \cite{FlormannAouane2017_SciReports}, which consists of a long-range attraction component and a
short-range repulsion component~\cite{Book_IntermolecularSurfaceForces},
and the combination of the two gives rise to adhesion at the separation
distance that minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. Such adhesive
potentials are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is essential for membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.

Adhesion between a vesicle and a substrate has been extensively studied,
primarily focused on analyzing the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase}
than the transient adhesion process; exceptions
include~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
%
% The short-range adhesive potential of a lipid bilayer membrane
% inevitably involves the membrane change redistribution and the membrane potential
% 
%  in relation the screening of charges in the electric double layer (EDL) right next to the
% membrane \cite{Book_IntermolecularSurfaceForces,LiraDimova}.
% When two vesicles are in close vicinity of each other within the electric double layer,  
% In this work we do not consider the electrostatic of two lipid bilayer membranes.
% 
Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ}, who measured the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker
constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds
to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion
energy per unit area to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes under adhesion in these regimes
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

>>>>

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by excess area (reduced volume), viscosity contrast
between interior and exterior fluids, and shear rate of the imposed
far-field fluid flow. In addition a vesicle with a rigid particle inside
is also investigated as a biological mimic of a eukaryotic cell with a
nucleus that occupies nearly half of the intracellular
volume~\cite{Veerapaneni2011_PRL}. Small-deformation analysis shows that
a vesicle tank-treads in a planar shear flow for low viscosity contrast
and shear rate. At high viscosity contrast the tank-tread (TT) dynamics
transitions to tumbling (TB)~\cite{Misbah2006_PRL, Vlahovska2007_PRE},
and this leads to a transition in effective shear viscosity in the
vesicle suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also
validated by direct numerical
simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,
ZabuskySegreDeschamps2011_PoF}.  Between TT and TB vesicle dynamics, a breathing (tremble) mode is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume  and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
Collective hydrodynamics of vesicles is dictated by the individual
vesicle-vesicle interactions.  
In a quiescent flow,
vesicle-vesicle adhesion leads to formation of vesicle doublets \cite{Ziherl2007_PRL,ZiherlSvetina2007_PNAS} or clusters \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
As a model for red blood cell (RBC) aggregates, an idealized model for adhesive vesicle-vesicle interactions gives rise to 
vesicle configurations similar to those observed in experiments of fibrinogen-induced RBC aggregates \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}. 
Using the L-J potential for point-point interaction between two vesicle membranes without the molecular details of
adhesive interactions between RBCs, Flormann {\it et al.} found the contact zone to buckle under strong adhesion \cite{FlormannAouane2017_SciReports}. 
A direct consequence of the buckling is the
sigmoidal shape of the adhesion region, which is also observed in RBC doublets and explained in a slightly different model \cite{ZiherlSvetina2007_PNAS}. 
It remains unknown how adhesive interactions between vesicles might lead to different vesicle hydrodynamics that results in
different rheological properties of a vesicle suspension. Such studies will provide useful insight to the rheology of blood flow and how to use nano solutes to control the hypertension
by adjusting the adhesive interactions between RBCs.
%

In numerical simulations of  a vesicle suspension, the adhesive interaction between two vesicles is often 
ignored \cite{Veerapaneni2009_JCP,RahimianVeerapaneniBiros2010_JCP} 
for numerical convenience, mostly 
because a smaller time step is often needed to resolve the dynamics of the lubrication thin film between two membranes with adhesion than without.
However, adhesive interactions  between RBCs lead to different RBC clusters that are expected to alter the rheological properties of the RBC suspension \cite{NeuMeiselman2002_BJ,SvetinaZiherl2008_Bioelectrochemistry}.
%
Hydrodynamic simulations of adhesive vesicles in a quiescent flow have revealed physical insights to the equilibrium shapes of RBC aggregates in both experiments 
\cite{FlormannAouane2017_SciReports} and theory \cite{ZiherlSvetina2007_PNAS}. The main goal in this work is to investigate the hydrodynamics of adhesive vesicles in flowing
conditions common in microfluidic vesicle flow, such as a planar shear flow or an extensional flow.

%
The electrostatic nature of lipid molecules (a hydrophobic tail and a hydrophilic head
with an electric dipole) complicates the interactions between a lipid
bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces,PerutkovaFrank-Bertoncelij2013_CSB} 
or a solid (such as a glass substrate
or a nano particle).  Adhesion between a vesicle and a substrate has been extensively studied,
primarily focused on analyzing the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase,SteinkuhlerAgudo-Canalejo2016_BJ}
than the transient adhesion process \cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
%
Adhesive interactions between lipid membranes are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is crucial to initiate membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.
In the absence of an external electric field and ions in the solvents, 
it is reported that 
%such complicated electrostatic component in 
the adhesive interactions between two membranes can be well approximated by the 
L-J-type potential \cite{FlormannAouane2017_SciReports}, which consists of a long-range attraction component and a
short-range repulsion component~\cite{Book_IntermolecularSurfaceForces},
and the combination of the two gives rise to adhesion at the separation
distance that minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. 
%%
%%move this somewhere else Jan30YNY
%Such adhesive
%potentials are essential in many biomedical, biological and biophysical
%processes.  For example, vesicle adhesion is essential for membrane
%fusion/fission in endocytosis and exocytosis, the transport of small
%vesicles through membrane surfaces.
%%move this somewhere else Jan30YNY

%Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ} used the micropipette to measure the adhesive force between two vesicles bound by strong adhesion.
They were able to  measure the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker
constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds
to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion
energy per unit area to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes under adhesion in these regimes
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

%Adhesion between a vesicle and a substrate has been extensively studied,
%primarily focused on analyzing the static equilibrium
%shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
%LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase}
%than the transient adhesion process; exceptions
%include~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
%
% The short-range adhesive potential of a lipid bilayer membrane
% inevitably involves the membrane change redistribution and the membrane potential
% 
%  in relation the screening of charges in the electric double layer (EDL) right next to the
% membrane \cite{Book_IntermolecularSurfaceForces,LiraDimova}.
% When two vesicles are in close vicinity of each other within the electric double layer,  
% In this work we do not consider the electrostatic of two lipid bilayer membranes.
%
